Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-09-18 paper/sections/proof-reconstruction.tex
fixed some references: \cite{Bertot2004}.

In the process to emulate \Metis' inference rules in type theory,
we encountered that some of \Metis' transformation for propositions in
the inference rules were given by \emph{general recursion} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.
Then, since a termination checker in type theory recognizes
\emph{structural recursion functions} as well-founded recursive
functions, we modify our algorithms in order to follow such a
definition mostly taken from \citeauthor{Abel2002}~\cite{Abel2002}.

\begin{definition} A recursive function is structurally recursive if
it calls itself with only \emph{structurally smaller} arguments.
\end{definition}

\begin{definition}
Given an \emph{ordering relation} between $t_1 : T$ and $t_2 : T$,
we say that $t_1$ is structurally smaller than $t_2$ if and only if,
USE \cite{Coquand1992} definition.
\begin{itemize}
  \item $t_1$ and $t_2$ are different constructor for $T$ type.
\end{itemize}
\end{definition}

To convert a general recursion function into a structural recursion
function, for the sake of simplicity, we will follow a technique
found mainly in \cite{Coquand1992,Abel2002,Bove2005}.
It mostly consists of using a \emph{recursor}, an inductive type, to
perform the \emph{induction} over its structure.
A type-checker will recognize structurally smaller arguments by using
the lexicographic order, as an extension of the structural ordering
by the recursor.

Let us consider the following example to exhibit the procedure
mentioned above for an inner function in the \strip inference.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt(φ)_{0} =
\begin{cases}
\unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\

@@

From the definition above, in the first case of the \unshunt
function, the size of the argument in the recursion,(φ₁ ∧ φ₂) ⇒ φ₃,
is not structured smaller than the decrease, (φ₁ ⇒ (φ₂ ⇒ φ₃)).
Therefore, a termination checking will fail with the \unshunt
function.

To improve that definition for terminating, we will use the
technique proposed by Bove and Capettra in
\cite{Bove2005}, which in particular with functions that
transform propositions consists of:
take a general recursive function $f: \Prop \to \Prop$.
and a predicate $P: \Prop \to \abbre{Ind}$, where $\abbre{Ind}$
is a type with a structured inductive definition.

Then, we redefine the function as
$f: \Prop \to \abbre{Ind}\ \to \Prop$, and perform the recursion
using induction on the structure of the \abbre{Ind} type.

In our case, the \abbre{Nat} data type for
natural numbers will be the $\abbre{Ind}$ type.
We define \abbre{Nat} as usual (constructors \zero and \suc).
We also allow common arithmetic operations and numbers represented by
$0,1,2,\dots$ that stand for $\zero, \suc, \suc(\zero), \dots$.

>>>>

In the process to emulate \Metis' inference rules in type theory,
we encountered that some of \Metis' transformation for propositions in
the inference rules were given by \emph{general recursion} functions.
But algorithms based on general recursions can not be translated
directly in type theory since it is not a guarantee they terminate.
Then, since a termination checker in type theory recognizes
\emph{structural recursion functions}, we modify our algorithms appropriately.

Using the concept of \emph{structurally smaller} from
\citeauthor{Abel2002}~\cite{Abel2002}. We say, a recursive function is
structurally recursive if it calls itself with only \emph{structurally smaller}
arguments.

To convert a general recursive call into a structurally recursive call, we use
the \emph{Bounded Recursion} technique presented in \cite{Coquand1992} but others found in
\cite{Coquand1992,Abel2002,Bove2005} could be used.

The \emph{Bounded Recursion} technique converts a general recursive
function into a structural recursive function by adding an additional
argument that handle the times the function can call itself.

In our case, we add a natural number $n : \Nat$ in the input of the general
recursive function for such an argument in mostly cases, and perform the
recursion over the structure of \Nat type. \abbre{Nat} type is the usual
type with \zero and \suc constructors, and including the common arithmetic
operations ($+,\,-,\,*$) and representations of $\zero, \suc, \suc(\zero),
\dots$
with $0,1,2,\dots$.

Then, the type-checker will
recognize structurally smaller arguments by using the lexicographic
order~\cite{Abel2002} for such kind of functions.

In other words, we take a general recursive function $f: \Prop \to \Prop$.
and we redefine that function as
$f: \Prop \to \Nat \to \Prop$, and perform the recursion
using induction on the structure of the \Nat type.

We define

Let us consider the following example to exhibit the technique
mentioned above for an inner function in the \strip inference.

\begin{equation}
\label{eq:unshunt-definition}
\unshunt_{0}(φ) =
\begin{cases}
\unshunt_{0}((φ₁ ∧ φ₂) ⇒ φ₃)
  & φ ≡ φ₁ ⇒ (φ₂ ⇒ φ₃),\\

@@

From the definition above, the size of the argument, ((φ₁ ∧ φ₂) ⇒ φ₃),
is not structurally smaller than the input formula, (φ₁ ⇒ (φ₂ ⇒ φ₃));
((φ₁ ∧ φ₂) ⇒ φ₃) is not a subformula of (φ₁ ⇒ (φ₂ ⇒ φ₃)).

### 2017-09-18 paper/sections/related-work.tex
[ paper ] working on structurally smaller arguments.
fixed some references: \cite{Bertot2004}.

\name{SMTCoq}~\cite{armand2011,Ekici2017} is a tool for the
\name{Coq} proof assistant \cite{coqteam} which provides a certified
checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit,
that calls \name{veriT} on any \name{Coq} goal.
In \cite{bezem2002automated},
given a fixed but arbitrary first-order signature, the authors
transform a proof produced by the first-order automatic theorem
prover \name{Bliksem} \cite{deNivelle2003} in a \name{Coq} proof
term.

>>>>

\name{SMTCoq}~\cite{armand2011,Ekici2017} is a tool for the
\name{Coq} proof assistant \cite{coqteam} which provides a certified
checker for proof witness coming from the \SMT solver
\name{veriT}~\cite{bouton2009} and adds a new tactic named verit,
that calls \name{veriT} on any \name{Coq} goal.
In \cite{Bezem2002},
given a fixed but arbitrary first-order signature, the authors
transform a proof produced by the first-order automatic theorem
prover \name{Bliksem} \cite{deNivelle2003} in a \name{Coq} proof
term.
